---
title: "Results"
---

This chapter presents the empirical results of applying bayesRVE's Three-Layer
Calibration to the @tannersmith2015 dataset, following the
experimental design detailed in @sec-tsl15-design. Results are organized in six
sections: CE calibration results, CHE calibration results, a combined CCC/CR2
ratio analysis with Gate 3 assessment, Satterthwaite degrees of freedom
comparison, variance component analysis, and an interim interpretation
summarizing the key findings.

```{r}
#| label: setup-ch12
#| echo: false
#| message: false
#| warning: false

library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
library(ggrepel)

R <- readRDS("../201_real-data-v060/v060_results.rds")

theme_brve <- theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "grey40", size = 11),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )

# Shared objects used across sections
dv_labs    <- R$table1$DV
K_dv       <- length(dv_labs)
ce_ratios  <- as.numeric(R$se_ratios_dv$CE_ratio)
che_ratios <- as.numeric(R$se_ratios_dv$CHE_ratio)
```


## MCMC Convergence {#sec-convergence}

All MCMC results in this chapter are based on converged posterior samples. Diagnostics confirm adequate mixing for both primary models: CE ($\hat{R}_{\max} = 1.008$, ESS$_{\text{bulk,min}} = 415$, ESS$_{\text{tail,min}} = 1{,}109$, divergences $= 0$) and CHE ($\hat{R}_{\max} = 1.005$, ESS$_{\text{bulk,min}} = 1{,}775$, ESS$_{\text{tail,min}} = 3{,}064$, divergences $= 0$). All parameters satisfy recommended thresholds ($\hat{R} < 1.01$, ESS $> 400$; Vehtari et al., 2021). Control-type models show similar diagnostics.


## CE Three-Layer Calibration Results {#sec-results}

The CE (Correlated Effects) model estimates a single between-study variance
component $\tau^2$. This section compares the frequentist CE baseline
(metafor + clubSandwich CR2) against the Bayesian CE model (bayesRVE + CCC).
The CE calibration pipeline was established in v0.5.0 and is unchanged in
v0.6.0; results here confirm backward compatibility and provide the baseline
for the CHE comparison that follows.


### CE Comparison Table

@tbl-ce-comparison presents the DV category estimates side by side. Point
estimates are nearly identical across methods, confirming that prior influence
on regression coefficients is negligible given $J = 107$ studies. The CCC SD
closely tracks the CR2 SE for all six categories.

```{r}
#| echo: false
#| label: tbl-ce-comparison
#| tbl-cap: "CE Model: metafor CR2 vs bayesRVE CCC-t (DV Category Moderators). Frequentist 'Est' = REML point estimate; Bayesian 'Est' = posterior mean. R_k is the CCC calibration ratio."

ce_cols <- c("DV", "CE_mf_est", "CE_mf_SE", "CE_mf_df",
             "bCE_est", "bCE_CCC_SD", "bCE_df", "bCE_Rk")
knitr::kable(
  R$table1[, ce_cols],
  digits = 3,
  col.names = c("DV Category", "Est (metafor)", "SE (CR2)", "df (metafor)",
                "Est (bayesRVE)", "CCC SD", "df (bayesRVE)", "$R_k$"),
  align = c("l", rep("r", 7))
)
```


### CE CCC/CR2 Ratio Table

@tbl-ce-ratio evaluates the primary validation metric: the ratio of Bayesian
CCC SD to frequentist CR2 SE. A ratio near 1.0 confirms that the CCC
posterior dispersion correctly targets the sandwich-based standard error.

```{r}
#| echo: false
#| label: tbl-ce-ratio
#| tbl-cap: "CE CCC SD / CR2 SE Ratio with Satterthwaite df Comparison"

ce_ratio_tbl <- R$se_ratios_dv[, c("DV", "CE_CCC_SD", "CE_CR2_SE", "CE_ratio",
                                     "CE_df_bayes", "CE_df_mf")]
ce_ratio_tbl$Gate3 <- ifelse(
  as.numeric(ce_ratio_tbl$CE_ratio) >= 0.9 &
    as.numeric(ce_ratio_tbl$CE_ratio) <= 1.1,
  "PASS", "FAIL"
)
knitr::kable(ce_ratio_tbl, digits = 4,
             col.names = c("DV Category", "CCC SD", "CR2 SE", "Ratio",
                           "df (bayesRVE)", "df (metafor)", "Gate 3"),
             align = c("l", rep("r", 5), "c"))
```

**Mean CE CCC/CR2 ratio:** `r sprintf("%.4f", mean(ce_ratios))` | **Range:** [`r sprintf("%.4f", min(ce_ratios))`, `r sprintf("%.4f", max(ce_ratios))`] | **Gate 3: `r if(all(ce_ratios >= 0.9 & ce_ratios <= 1.1)) "ALL PASS" else "CHECK REQUIRED"`**

The tight clustering of CE ratios around 1.0 confirms that Cholesky
Calibration Correction achieves its design objective: the Bayesian CCC
posterior spread matches the frequentist CR2 sandwich standard error to
within a few percent for all well-estimated DV category coefficients.


### CE Posterior Density Overlay

@fig-ce-posterior-overlay displays the original Bayesian posterior (blue)
alongside the CCC-calibrated posterior (red) for each DV category. The
near-complete overlap confirms that CCC acts as a near-identity transformation
for well-estimated coefficients with high effective degrees of freedom.

```{r}
#| echo: false
#| label: fig-ce-posterior-overlay
#| fig-cap: "CE beta posterior overlay: original (blue) vs CCC-calibrated (red) for 6 DV categories. Near-identical densities indicate CCC acts as a near-identity mapping when the posterior already approximates the CR2 sandwich variance."
#| fig-width: 10
#| fig-height: 7

draws_orig_ce <- R$cal_ce_dv$draws_orig[, 1:K_dv]
draws_star_ce <- R$cal_ce_dv$draws_star[, 1:K_dv]

overlay_ce <- bind_rows(
  lapply(1:K_dv, function(k) {
    data.frame(
      value = c(draws_orig_ce[, k], draws_star_ce[, k]),
      Type  = rep(c("Original", "CCC-calibrated"), each = nrow(draws_orig_ce)),
      DV    = dv_labs[k],
      stringsAsFactors = FALSE
    )
  })
) |>
  mutate(DV = factor(DV, levels = dv_labs))

ggplot(overlay_ce, aes(x = value, fill = Type, color = Type)) +
  geom_density(alpha = 0.4, linewidth = 0.6) +
  facet_wrap(~ DV, scales = "free", ncol = 3) +
  scale_fill_manual(values = c("Original" = "#4393C3",
                                "CCC-calibrated" = "#D6604D")) +
  scale_color_manual(values = c("Original" = "#2166AC",
                                 "CCC-calibrated" = "#B2182B")) +
  labs(x = "Coefficient Estimate (g)", y = "Density",
       title = "CE Posterior Densities: Original vs CCC-Calibrated",
       subtitle = "Near-complete overlap indicates R_k close to 1.0") +
  theme_brve +
  theme(legend.title = element_blank())
```


### CE Tau Posterior with REML Reference

@fig-ce-tau shows the posterior distribution of $\tau$ (between-study SD)
under the CE model, with the metafor REML point estimate marked for reference.
The Bayesian posterior mean is systematically larger than REML, reflecting
the half-Cauchy(0, 0.5) prior's heavier tail.

```{r}
#| echo: false
#| label: fig-ce-tau
#| fig-cap: "CE tau posterior density. The dashed vertical line marks the metafor REML estimate. The Bayesian posterior mean exceeds REML due to the half-Cauchy(0, 0.5) prior, which places more mass on larger values than the constrained REML optimizer."
#| fig-width: 7
#| fig-height: 4.5

tau_df_ce <- data.frame(tau = R$tau_ce_dv)
tau_reml_ce <- R$vc_table$tau_est[R$vc_table$Model == "CE (metafor)"]

ggplot(tau_df_ce, aes(x = tau)) +
  geom_density(fill = "#4393C3", alpha = 0.5, color = "#2166AC",
               linewidth = 0.7) +
  geom_vline(xintercept = tau_reml_ce, linetype = "dashed",
             color = "#B2182B", linewidth = 0.9) +
  annotate("text", x = tau_reml_ce + 0.01, y = Inf, vjust = 2, hjust = 0,
           label = sprintf("REML = %.3f", tau_reml_ce),
           color = "#B2182B", size = 4, fontface = "italic") +
  labs(x = expression(tau ~ "(between-study SD)"), y = "Density",
       title = expression("CE" ~ tau ~ "Posterior Distribution"),
       subtitle = "Half-Cauchy(0, 0.5) prior; dashed line = metafor REML") +
  theme_brve
```


## CHE Three-Layer Calibration Results {#sec-results-che}

The CHE (Correlated-Hierarchical Effects) model decomposes heterogeneity
into between-study ($\tau^2$) and within-study ($\omega^2$) components.
Extending Three-Layer Calibration to the CHE model is the central
contribution of bayesRVE v0.6.0. This section presents the CHE calibration
results, including the tau-omega joint posterior that reveals the
decomposition structure.


### CHE Comparison Table

@tbl-che-comparison presents the CHE model estimates. Point estimates
remain concordant across frequentist and Bayesian approaches. CCC SD
tracks CR2 SE closely, confirming that the omega-aware CCC machinery
(Decision D20: Jensen-safe $\hat\omega^2$; Decision D24: Woodbury
$W^{1/2}$) produces calibrated inference.

```{r}
#| echo: false
#| label: tbl-che-comparison
#| tbl-cap: "CHE Model: metafor CR2 vs bayesRVE CCC-t (DV Category Moderators). Frequentist 'Est' = REML; Bayesian 'Est' = posterior mean. CHE CCC uses omega-aware Woodbury weights (v0.6.0)."

che_cols <- c("DV", "CHE_mf_est", "CHE_mf_SE", "CHE_mf_df",
              "bCHE_est", "bCHE_CCC_SD", "bCHE_df", "bCHE_Rk")
knitr::kable(
  R$table1[, che_cols],
  digits = 3,
  col.names = c("DV Category", "Est (metafor)", "SE (CR2)", "df (metafor)",
                "Est (bayesRVE)", "CCC SD", "df (bayesRVE)", "$R_k$"),
  align = c("l", rep("r", 7))
)
```


### CHE CCC/CR2 Ratio Table with Gate 3 Assessment

@tbl-che-ratio is the Gate 3 evaluation table. Each DV category's CCC/CR2
ratio is assessed against the pre-specified [0.9, 1.1] acceptance band.

```{r}
#| echo: false
#| label: tbl-che-ratio
#| tbl-cap: "CHE CCC SD / CR2 SE Ratio and Satterthwaite df (Gate 3 Assessment)"

che_ratio_tbl <- R$se_ratios_dv[, c("DV", "CHE_CCC_SD", "CHE_CR2_SE",
                                      "CHE_ratio", "CHE_df_bayes", "CHE_df_mf")]
che_ratio_tbl$Gate3 <- ifelse(
  as.numeric(che_ratio_tbl$CHE_ratio) >= 0.9 &
    as.numeric(che_ratio_tbl$CHE_ratio) <= 1.1,
  "PASS", "FAIL"
)
knitr::kable(che_ratio_tbl, digits = 4,
             col.names = c("DV Category", "CCC SD", "CR2 SE", "Ratio",
                           "df (bayesRVE)", "df (metafor)", "Gate 3"),
             align = c("l", rep("r", 5), "c"))
```

```{r}
#| echo: false
mean_che <- mean(che_ratios)
all_che_pass <- all(che_ratios >= 0.9 & che_ratios <= 1.1)
```

**Mean CHE CCC/CR2 ratio:** `r sprintf("%.4f", mean_che)` | **Range:** [`r sprintf("%.4f", min(che_ratios))`, `r sprintf("%.4f", max(che_ratios))`]

**Gate 3 verdict: `r if(all_che_pass) "ALL PASS" else "CHECK REQUIRED"`** --- All 6
DV categories fall within the [0.9, 1.1] acceptance band, confirming that
the CHE extension of CCC produces calibrated inference. The Bayesian CCC
posterior spread matches the frequentist CR2 sandwich standard error for
the CHE model, validating the omega-aware weight construction (Decisions
D20 and D24).


### CHE Posterior Density Overlay

```{r}
#| echo: false
#| label: fig-che-posterior-overlay
#| fig-cap: "CHE beta posterior overlay: original (blue) vs CCC-calibrated (red) for 6 DV categories. Near-identical densities confirm that CCC acts as a near-identity mapping for the CHE model when coefficients are well-estimated with sufficient clusters."
#| fig-width: 10
#| fig-height: 7

draws_orig_che <- R$cal_che_dv$draws_orig[, 1:K_dv]
draws_star_che <- R$cal_che_dv$draws_star[, 1:K_dv]

overlay_che <- bind_rows(
  lapply(1:K_dv, function(k) {
    data.frame(
      value = c(draws_orig_che[, k], draws_star_che[, k]),
      Type  = rep(c("Original", "CCC-calibrated"), each = nrow(draws_orig_che)),
      DV    = dv_labs[k],
      stringsAsFactors = FALSE
    )
  })
) |>
  mutate(DV = factor(DV, levels = dv_labs))

ggplot(overlay_che, aes(x = value, fill = Type, color = Type)) +
  geom_density(alpha = 0.4, linewidth = 0.6) +
  facet_wrap(~ DV, scales = "free", ncol = 3) +
  scale_fill_manual(values = c("Original" = "#4393C3",
                                "CCC-calibrated" = "#D6604D")) +
  scale_color_manual(values = c("Original" = "#2166AC",
                                 "CCC-calibrated" = "#B2182B")) +
  labs(x = "Coefficient Estimate (g)", y = "Density",
       title = "CHE Posterior Densities: Original vs CCC-Calibrated",
       subtitle = "Near-complete overlap indicates CCC near-identity for well-estimated coefficients") +
  theme_brve +
  theme(legend.title = element_blank())
```


### Tau-Omega Joint Posterior

@fig-joint-posterior is the signature figure for the CHE analysis. The joint
posterior of $\tau$ and $\omega$ reveals the decomposition structure, the
relationship between variance components, and how the Bayesian estimates
relate to the REML point estimates.

```{r}
#| echo: false
#| label: fig-joint-posterior
#| fig-cap: "Joint posterior distribution of tau (between-study SD) and omega (within-study SD) from the CHE DV model. Black contours show posterior density levels. The red cross marks the REML point estimate from metafor. The Bayesian posterior for omega is substantially shrunk toward zero relative to REML, reflecting prior regularization of the within-study component."
#| fig-width: 8
#| fig-height: 7

joint_df <- tibble(
  tau   = R$tau_che_dv,
  omega = R$omega_che_dv
)

tau_reml_che   <- as.numeric(R$vc_table$tau_est[R$vc_table$Model == "CHE (metafor)"])
omega_reml_che <- as.numeric(R$vc_table$omega_est[R$vc_table$Model == "CHE (metafor)"])

ggplot(joint_df, aes(x = tau, y = omega)) +
  geom_point(alpha = 0.03, size = 0.3, color = "grey40") +
  stat_density_2d(color = "black", linewidth = 0.5) +
  annotate("point", x = tau_reml_che, y = omega_reml_che,
           shape = 4, size = 5, stroke = 2, color = "#B2182B") +
  annotate("text", x = tau_reml_che + 0.015, y = omega_reml_che + 0.008,
           label = "REML", color = "#B2182B", fontface = "bold", size = 4) +
  labs(title = expression("Joint Posterior:" ~ tau ~ "vs" ~ omega ~
                          "(CHE DV Model)"),
       subtitle = "Red cross = metafor REML estimate; contours = posterior density",
       x = expression(tau ~ "(between-study SD)"),
       y = expression(omega ~ "(within-study SD)")) +
  theme_brve
```

Three features of the joint posterior merit attention. First, the posterior
mass for $\tau$ is centered around `r sprintf("%.3f", median(R$tau_che_dv))`,
substantially higher than the REML estimate of 0.123, reflecting the
half-Cauchy prior's heavier tail. Second, the posterior for $\omega$ is
centered around `r sprintf("%.3f", median(R$omega_che_dv))`, well below the
REML estimate of 0.116, because the half-Cauchy prior induces stronger
shrinkage for the within-study component, which is harder to identify with
limited per-study replications. Third, the posterior correlation between $\tau$ and $\omega$ is near zero ($r = -0.036$), indicating that partial confounding between variance components is minimal in this well-identified dataset ($J = 107$). The slight negative sign is directionally consistent with the theoretical trade-off between between-study and within-study heterogeneity.


### Marginal Posteriors: Tau and Omega

@fig-marginal-posteriors provides a clearer view of the individual variance
component distributions. The left panel compares $\tau$ posteriors under CE
and CHE, showing how the CE model absorbs all heterogeneity into a single
component. The right panel shows the CHE-specific $\omega$ posterior.

```{r}
#| echo: false
#| label: fig-marginal-posteriors
#| fig-cap: "Marginal posterior densities. Left: tau for CE (blue) and CHE (red), with REML reference lines. Right: omega for CHE only, with REML reference. The CE model absorbs all heterogeneity into tau; CHE partitions it, yielding a smaller tau and a separate omega."
#| fig-width: 11
#| fig-height: 5

tau_comp_df <- bind_rows(
  tibble(value = R$tau_ce_dv,  Model = "CE"),
  tibble(value = R$tau_che_dv, Model = "CHE")
)

p_tau <- ggplot(tau_comp_df, aes(x = value, fill = Model)) +
  geom_density(alpha = 0.4, color = NA) +
  geom_vline(xintercept = tau_reml_ce, linetype = "dashed",
             color = "#4393C3", linewidth = 0.8) +
  geom_vline(xintercept = tau_reml_che, linetype = "dashed",
             color = "#D6604D", linewidth = 0.8) +
  scale_fill_manual(values = c("CE" = "#4393C3", "CHE" = "#D6604D")) +
  annotate("text", x = tau_reml_ce + 0.005, y = Inf, vjust = 2,
           label = "CE REML", color = "#4393C3", size = 3.5,
           fontface = "italic") +
  annotate("text", x = tau_reml_che - 0.005, y = Inf, vjust = 3.5,
           label = "CHE REML", color = "#D6604D", size = 3.5,
           fontface = "italic", hjust = 1) +
  labs(title = expression("Marginal Posterior:" ~ tau),
       x = expression(tau ~ "(between-study SD)"), y = "Density",
       fill = NULL) +
  theme_brve + theme(legend.position = "top")

omega_df <- tibble(value = R$omega_che_dv)

p_omega <- ggplot(omega_df, aes(x = value)) +
  geom_density(fill = "#D6604D", alpha = 0.4, color = NA) +
  geom_vline(xintercept = omega_reml_che, linetype = "dashed",
             color = "#B2182B", linewidth = 0.8) +
  annotate("text", x = omega_reml_che + 0.01, y = Inf, vjust = 2,
           label = "CHE REML", color = "#B2182B", size = 3.5,
           fontface = "italic") +
  labs(title = expression("Marginal Posterior:" ~ omega),
       x = expression(omega ~ "(within-study SD)"), y = "Density") +
  theme_brve

p_tau + p_omega + plot_layout(widths = c(1, 1))
```

The CE model has no $\omega$ parameter: all heterogeneity is absorbed
into $\tau$, producing $\tau_\text{CE} > \tau_\text{CHE}$. The CHE model
partitions heterogeneity into two components, yielding a smaller
$\tau_\text{CHE}$ and a separate $\omega > 0$. The Bayesian $\omega$
posterior (median `r sprintf("%.3f", median(R$omega_che_dv))`) is
substantially below the REML estimate (0.116), reflecting prior shrinkage
that is characteristic of half-Cauchy priors on variance components near
the identification boundary.


## CCC/CR2 Ratio Analysis {#sec-results-ratio}

This section consolidates the CE and CHE ratio analyses into a unified
assessment. The combined view provides the definitive Gate 3 evaluation
for bayesRVE v0.6.0 on the TSL15 dataset.


### Combined Ratio Bar Chart with Gate 3 Band

@fig-ratio-bars displays all 12 CCC/CR2 ratios (6 DV categories $\times$
2 models) with the Gate 3 acceptance band. This is the primary validation
figure for the technical report.

```{r}
#| echo: false
#| label: fig-ratio-bars
#| fig-cap: "CCC SD / CR2 SE ratio for each DV category, separately for CE (blue) and CHE (red). Segments show deviation from parity (1.0). The green band marks the Gate 3 acceptance range [0.9, 1.1]. All 12 ratios fall within the band, confirming calibration success for both models."
#| fig-width: 10
#| fig-height: 6

ratio_df <- bind_rows(
  data.frame(DV = dv_labs, Ratio = ce_ratios,
             Model = "CE", stringsAsFactors = FALSE),
  data.frame(DV = dv_labs, Ratio = che_ratios,
             Model = "CHE", stringsAsFactors = FALSE)
) |>
  mutate(DV = factor(DV, levels = rev(dv_labs)))

# Manual dodge: CE offset -0.15, CHE offset +0.15
ratio_df <- ratio_df |>
  mutate(y_num = as.numeric(DV) + ifelse(Model == "CE", -0.15, 0.15))

ggplot(ratio_df, aes(colour = Model)) +
  annotate("rect", xmin = 0.9, xmax = 1.1, ymin = -Inf, ymax = Inf,
           fill = "#2CA02C", alpha = 0.1) +
  geom_vline(xintercept = 1.0, linetype = "solid", color = "grey40") +
  geom_segment(aes(x = 1.0, xend = Ratio, y = y_num, yend = y_num),
               linewidth = 0.8, show.legend = FALSE) +
  geom_point(aes(x = Ratio, y = y_num), size = 3) +
  scale_colour_manual(values = c("CE" = "#4393C3", "CHE" = "#D6604D")) +
  scale_y_continuous(breaks = seq_along(levels(ratio_df$DV)),
                     labels = levels(ratio_df$DV)) +
  coord_cartesian(xlim = c(0.85, 1.15)) +
  scale_x_continuous(breaks = seq(0.85, 1.15, 0.05)) +
  labs(title = "CCC SD / CR2 SE Ratio by DV Category",
       subtitle = "Green band = Gate 3 acceptance range [0.9, 1.1]",
       x = "CCC SD / CR2 SE Ratio", y = NULL) +
  theme_brve
```


### SE Scatter: CCC SD vs CR2 SE

@fig-se-scatter provides a complementary view: each point represents a
DV category, with CR2 SE on the horizontal axis and CCC SD on the vertical
axis. Points near the dashed diagonal indicate agreement.

```{r}
#| echo: false
#| label: fig-se-scatter
#| fig-cap: "CCC SD vs CR2 SE for each DV category, faceted by model. Points near the dashed diagonal indicate agreement between Bayesian CCC and frequentist CR2. Labels identify each DV category."
#| fig-width: 11
#| fig-height: 6

se_scatter <- bind_rows(
  data.frame(CR2_SE = as.numeric(R$se_ratios_dv$CE_CR2_SE),
             CCC_SD = as.numeric(R$se_ratios_dv$CE_CCC_SD),
             Model = "CE", DV = dv_labs, stringsAsFactors = FALSE),
  data.frame(CR2_SE = as.numeric(R$se_ratios_dv$CHE_CR2_SE),
             CCC_SD = as.numeric(R$se_ratios_dv$CHE_CCC_SD),
             Model = "CHE", DV = dv_labs, stringsAsFactors = FALSE)
)

se_range <- range(c(se_scatter$CR2_SE, se_scatter$CCC_SD)) * c(0.9, 1.1)

ggplot(se_scatter, aes(x = CR2_SE, y = CCC_SD, label = DV)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey50") +
  geom_point(aes(color = Model), size = 3.5) +
  geom_text_repel(size = 3, max.overlaps = 20, show.legend = FALSE) +
  scale_color_manual(values = c("CE" = "#4393C3", "CHE" = "#D6604D")) +
  coord_equal(xlim = se_range, ylim = se_range) +
  facet_wrap(~ Model, ncol = 2) +
  labs(title = "CCC SD vs CR2 SE (DV Category Model)",
       subtitle = "Dashed line = perfect agreement (ratio = 1.0)",
       x = "Frequentist CR2 SE", y = "Bayesian CCC SD") +
  theme_brve +
  theme(legend.position = "none")
```


### Grand Summary Table

@tbl-grand-summary consolidates the key validation metrics for both models.

```{r}
#| echo: false
#| label: tbl-grand-summary
#| tbl-cap: "Grand Calibration Summary: CE and CHE Models on TSL15 DV Category Specification"

ce_pass  <- all(ce_ratios >= 0.9 & ce_ratios <= 1.1)
che_pass <- all(che_ratios >= 0.9 & che_ratios <= 1.1)

grand_df <- data.frame(
  Metric = c("Mean CCC/CR2 ratio",
             "Ratio range",
             "Gate 3 [0.9, 1.1]"),
  CE = c(
    sprintf("%.4f", mean(ce_ratios)),
    sprintf("[%.4f, %.4f]", min(ce_ratios), max(ce_ratios)),
    if (ce_pass) "PASS" else "FAIL"
  ),
  CHE = c(
    sprintf("%.4f", mean(che_ratios)),
    sprintf("[%.4f, %.4f]", min(che_ratios), max(che_ratios)),
    if (che_pass) "PASS" else "FAIL"
  ),
  stringsAsFactors = FALSE
)

knitr::kable(grand_df, col.names = c("Metric", "CE", "CHE"),
             align = c("l", "r", "r"))
```

Both models pass Gate 3 for all 6 DV categories. The CE ratios cluster
tightly around 1.0, consistent with the mature CE calibration pipeline.
The CHE ratios show slightly wider dispersion, consistent with the
additional variance component estimation introducing modest MCMC variability,
but all remain within the acceptance band.


## Satterthwaite df Comparison {#sec-results-df}

Satterthwaite degrees of freedom define the reference distribution for
Three-Layer Calibration's Layer 3 ($t$-based inference). This section
examines whether bayesRVE's independently-implemented Satterthwaite df
(Decision D11: no clubSandwich runtime dependency) produces consistent
results with the frequentist implementation.


### df Scatter: bayesRVE vs metafor

@fig-df-scatter plots the Satterthwaite df from bayesRVE against those from
metafor + clubSandwich for each DV category under both models.

```{r}
#| echo: false
#| label: fig-df-scatter
#| fig-cap: "Satterthwaite df: bayesRVE vs metafor for CE (blue) and CHE (red). Points above the diagonal indicate bayesRVE reports higher df; below indicates lower. Systematic deviations arise from differences in plug-in variance components (posterior mean vs REML)."
#| fig-width: 8
#| fig-height: 7

df_scatter <- bind_rows(
  data.frame(DV = dv_labs,
             df_bayes = as.numeric(R$se_ratios_dv$CE_df_bayes),
             df_freq  = as.numeric(R$se_ratios_dv$CE_df_mf),
             Model = "CE", stringsAsFactors = FALSE),
  data.frame(DV = dv_labs,
             df_bayes = as.numeric(R$se_ratios_dv$CHE_df_bayes),
             df_freq  = as.numeric(R$se_ratios_dv$CHE_df_mf),
             Model = "CHE", stringsAsFactors = FALSE)
)

df_range <- range(c(df_scatter$df_bayes, df_scatter$df_freq)) * c(0.8, 1.1)

ggplot(df_scatter, aes(x = df_freq, y = df_bayes, color = Model,
                        label = DV)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey50") +
  geom_point(size = 3.5) +
  geom_text_repel(size = 3, max.overlaps = 20, show.legend = FALSE) +
  scale_color_manual(values = c("CE" = "#4393C3", "CHE" = "#D6604D")) +
  coord_equal(xlim = df_range, ylim = df_range) +
  labs(title = "Satterthwaite df: bayesRVE vs metafor",
       subtitle = "Per-coefficient df for DV category moderators",
       x = "metafor + clubSandwich df", y = "bayesRVE df") +
  theme_brve
```


### Why Degrees of Freedom Differ

The Satterthwaite df depends on the Hat matrix
$H = X(X'WX)^{-1}X'W$, which in turn depends on the working weights $W$.
bayesRVE and metafor use different plug-in variance component estimates to
construct $W$, and these differences propagate through the entire
$W \to H \to \text{CR2 adjustment} \to \text{Satterthwaite df}$ pipeline.

@tbl-vc-plugin documents the plug-in variance components used by each
method for the DV category model.

```{r}
#| echo: false
#| label: tbl-vc-plugin
#| tbl-cap: "Plug-in Variance Components Used in Weight Construction (DV Category Model)"

vc_plugin <- data.frame(
  Component = c("tau (CE)", "tau (CHE)", "omega (CHE)"),
  metafor_REML = c(
    sprintf("%.3f", as.numeric(R$vc_table$tau_est[R$vc_table$Model == "CE (metafor)"])),
    sprintf("%.3f", as.numeric(R$vc_table$tau_est[R$vc_table$Model == "CHE (metafor)"])),
    sprintf("%.3f", as.numeric(R$vc_table$omega_est[R$vc_table$Model == "CHE (metafor)"]))
  ),
  bayesRVE_post = c(
    sprintf("%.3f", mean(R$tau_ce_dv)),
    sprintf("%.3f", mean(R$tau_che_dv)),
    sprintf("%.3f", sqrt(mean(R$omega_che_dv^2)))
  ),
  stringsAsFactors = FALSE
)

knitr::kable(vc_plugin,
             col.names = c("Component", "metafor (REML)",
                           "bayesRVE (posterior mean)"),
             align = c("l", "r", "r"))
```

The larger Bayesian $\hat\tau$ increases the relative contribution of
between-study variance to the working model, altering study-level leverage
values and thus the effective degrees of freedom. The smaller Bayesian
$\hat\omega$ (prior shrinkage) further modifies the CHE weight structure.
These differences are expected and do not indicate a calibration failure:
the df discrepancy is a natural consequence of using Bayesian rather than
REML point estimates for plug-in quantities, and it is exactly this
difference that Layer 3 of the calibration pipeline is designed to handle.

The key observation is that the *direction* of the CCC SD adjustment (via
$L_2$) and the *magnitude* of the df adjustment (via the Satterthwaite
formula) are jointly determined by the same plug-in variance components.
When $\hat\tau_\text{Bayes} > \hat\tau_\text{REML}$, both the sandwich
variance and the degrees of freedom shift in compensating directions,
maintaining the overall calibration of the CCC-$t$ interval.


## Variance Components {#sec-results-vc}

Understanding the variance component estimates is essential for interpreting
both the CCC mechanism and the substantive findings. This section compares
$\tau$ and $\omega$ estimates across all five estimators.


### Variance Component Comparison Table

@tbl-vc-all presents the variance component estimates (on the SD scale) for
all models fitted to the DV category specification.

```{r}
#| echo: false
#| label: tbl-vc-all
#| tbl-cap: "Variance Component Estimates (SD Scale) Across All Models, DV Category Specification"

knitr::kable(R$vc_table, digits = 4,
             col.names = c("Model", "tau (SD)", "omega (SD)"),
             align = c("l", "r", "r"))
```


### Variance Components Bar Chart

@fig-vc-bars visualizes the variance component structure across models,
illustrating the tau-absorption phenomenon in CE models and the
decomposition in CHE models.

```{r}
#| echo: false
#| label: fig-vc-bars
#| fig-cap: "Variance component estimates (SD scale) across all models. CE models estimate only tau (which absorbs omega). CHE models decompose heterogeneity into between-study (tau) and within-study (omega) components. Bayesian tau estimates exceed REML due to prior influence; Bayesian omega is shrunk below REML."
#| fig-width: 9
#| fig-height: 5.5

vc_long <- R$vc_table |>
  pivot_longer(cols = c(tau_est, omega_est),
               names_to = "Component", values_to = "Estimate") |>
  filter(!is.na(Estimate)) |>
  mutate(
    Component = ifelse(Component == "tau_est", "tau", "omega"),
    Model = factor(Model, levels = R$vc_table$Model)
  )

ggplot(vc_long, aes(x = Model, y = Estimate, fill = Component)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6,
           alpha = 0.85) +
  scale_fill_manual(values = c("tau" = "#4393C3", "omega" = "#D6604D"),
                    labels = c(expression(tau), expression(omega))) +
  labs(x = NULL, y = "Estimate (SD scale)",
       title = "Variance Components by Model",
       subtitle = "CE absorbs omega into tau; CHE decomposes both") +
  theme_brve +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 10))
```


### Interpretation

The variance component comparison reveals three systematic patterns.

**Tau absorption in CE models.** The CE model has a single between-study
random effect ($\tau^2$) that must account for all sources of heterogeneity
beyond sampling error, including both genuine between-study variation and
within-study variation across dependent effect sizes. As a result,
$\tau_\text{CE} > \tau_\text{CHE}$ for both frequentist and Bayesian
estimators. The magnitude of this inflation depends on the true $\omega^2$:
when within-study heterogeneity is substantial, the CE $\tau$ estimate
absorbs a correspondingly larger share.

**Bayesian tau exceeds REML.** Across both CE and CHE models, the Bayesian
posterior mean of $\tau$ exceeds the REML point estimate. This reflects the
half-Cauchy(0, 0.5) prior, which has heavier tails than the implicit REML
constraint (which truncates the likelihood at the boundary $\tau^2 = 0$).
The half-Cauchy prior places more mass on larger values of $\tau$, pulling
the posterior mean upward relative to REML, especially when the data provide
moderate (rather than overwhelming) information about the between-study
variance.

**Bayesian omega shrunk below REML.** The CHE Bayesian $\omega$ posterior
is centered below the REML estimate. The within-study variance component
is harder to identify than the between-study component because its
estimation relies on within-study replication, which varies substantially
across studies. The half-Cauchy prior induces more pronounced shrinkage
for $\omega$ than for $\tau$ because the data provide less information to
overcome the prior's pull toward zero.


## Interim Interpretation {#sec-results-interpretation}

The empirical results from the TSL15 DV category specification yield six
key findings.

1. **CCC/CR2 ratio near unity for both models.** The mean ratio is
   `r sprintf("%.4f", mean(ce_ratios))` for CE and
   `r sprintf("%.4f", mean(che_ratios))` for CHE, with all 12 individual
   ratios falling within the pre-specified [0.9, 1.1] acceptance band.
   This confirms that Cholesky Calibration Correction successfully maps
   the Bayesian posterior to have the CR2 sandwich variance as its
   dispersion target.

2. **Gate 3: PASS.** Both CE and CHE models pass the pre-specified Gate 3
   criterion on all 6 DV category coefficients. The CHE result validates
   the v0.6.0 extension of CCC to models with two variance components.
   Gate 3 does not apply to covariate coefficients, whose CCC/CR2 ratios
   may exceed the [0.9, 1.1] band due to lower effective cluster counts
   (see @sec-mechanics-covariates for details).

3. **Point estimate concordance.** Bayesian posterior means and frequentist
   REML estimates are nearly identical for all regression coefficients,
   confirming that prior influence on location parameters is negligible
   given the large sample ($J = 107$, $N = 1{,}198$).

4. **Satterthwaite df differ systematically.** bayesRVE and metafor produce
   different Satterthwaite degrees of freedom because they use different
   plug-in variance components ($\hat\tau_\text{Bayes} > \hat\tau_\text{REML}$;
   $\hat\omega_\text{Bayes} < \hat\omega_\text{REML}$). This is expected
   and does not compromise calibration: the CCC SD and df adjustments are
   jointly determined and compensate for each other.

5. **Tau-omega decomposition reveals prior shrinkage structure.** The CHE
   joint posterior shows near-zero correlation between $\tau$ and $\omega$
   ($r = -0.036$), indicating minimal confounding in this well-identified
   dataset, with $\omega$ substantially shrunk below REML. This prior shrinkage is
   scientifically meaningful: it expresses genuine uncertainty about whether
   within-study heterogeneity is present, given that $\omega$ is harder to
   identify than $\tau$.

6. **CCC as near-identity for well-estimated parameters.** For the 6 DV
   category coefficients (high effective df, many contributing clusters),
   the CCC calibration ratio $R_k$ is close to 1.0 in both models. CCC
   has minimal work to do when the Bayesian posterior already approximates
   the CR2 sandwich variance.

These findings establish the empirical foundation for Three-Layer
Calibration. The next chapter (@sec-mechanics) examines the CCC mechanism
in greater depth, analyzing the $R_k$ gradient, the multivariate
transformation structure, and the three-layer interval width decomposition
that reveals how each calibration layer contributes to the final inference.
